On the presentation of orthogonal groups over polynomial rings

### 2019-01-08 dl-paper.tex
Added more formulas

\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark} %\newtheorem{rem}[lemma]{Remark} \Crefname{rem}{Remark}{Remarks}

@@

Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$

@@

Recall the definition of the Dennis--Stein symbol (cf. III.5.11 within Weibel's book).
\[ \langle r,s \rangle _ \alpha = x_{-\alpha}\left(-\frac{s}{1 - rs}\right) \cdot x_{\alpha}(-r) \cdot x_{-\alpha}(s) \cdot x_{\alpha}\left(\frac{r}{1-rs}\right) \cdot h_{\alpha}(1 - rs)^{-1}, \]
where $r, s$ are arbitrary elements of $R$ such that $1 - rs\in R^*$ then

@@

For any $\alpha \in \Phi$ one can find $\beta$ such that $\langle \alpha, \beta \rangle = -1$
(except for the case when $\alpha$ is a long root of a symplectic root system).
Choose $\beta$ such that $\langle \alpha, \beta \rangle = -1$
\begin{multline}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(X m) %= x_\alpha(X^{\langle \alpha, \beta \rangle}a) x_{-\alpha}(X^{-\langle \alpha, \beta \rangle}m) 
 = {}^{h_\beta(X)}(x_\alpha(a) x_{-\alpha}(m)) = \\
 = {}^{h_\beta(X)}\left( x_{-\alpha}\left(\frac{m}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\frac{a}{1+am}\right) \right) = \\
 x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha(X^{-1})^{-1} x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot \{1+am, X^{-1}\}^{-1} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}
\end{multline}

We need to introduce notation for certain subgroups of $\St(A[X, X^{-1}])$:
\begin{align}
 G_+   & = \mathrm{Im}(\St(\Phi, A[X], Xm[X]) \to \St(A[X, X^{-1}]))\\
 G_+^0 & = \mathrm{Im}(\St(\Phi, A[X], m[X]) \to \St(A[X, X^{-1}]))
\end{align}
Notice that for $ g\in G_+$ the element $ev_{X=0}(g)^{-1}g$ lies in $G_+^0$.
\begin{lemma} Assume that $\Phi$ is simply-laced.
 For any $a\in A$, $\alpha\in \Phi$, $g_0 \in G^0_+$ there exist $m'\in m$ and $g'\in G_+$ such that 
 \[x_{\alpha}(aX^{-1}) g_0  = g' x_\alpha(-a(1+m')X^{-1}) \cdot \{X, 1+m'\}.\]
\end{lemma}
\begin{proof}
 Choose some parabolic subset $S$ of roots in $\Phi$ in such a way that its special part $\Sigma(S)$ does not contain the root $\alpha$.
 In view of~\cref{lem:Zgen} the subgroup $G_+^0$ is generated by images in $\St(\Phi, A[X, X^{-1}]$ of the elements of the set $\mathcal{Z}(\Sigma(S), A[X], Xm[X])$.

>>>>

\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}

@@

\begin{comment}
Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$ 
\end{comment}

@@

Recall the definition of the Dennis--Stein symbol (cf. III.5.11 within Weibel's book).
\begin{equation} \label{eq:dennis-stein}
 \langle r,s \rangle _ \alpha = x_{-\alpha}\left(-\frac{s}{1 - rs}\right) \cdot x_{\alpha}(-r) \cdot x_{-\alpha}(s) \cdot x_{\alpha}\left(\frac{r}{1-rs}\right) \cdot h_{\alpha}(1 - rs)^{-1},
\end{equation} 
where $r, s$ are arbitrary elements of $R$ such that $1 - rs\in R^*$ then

@@

We need to introduce notation for certain subgroups of $\St(A[X, X^{-1}])$:
\begin{align}
 G_+^0 & = \mathrm{Im}(\St(\Phi, A[X], XM[X]) \to \St(A[X, X^{-1}]))\\
 G_+   & = \mathrm{Im}(\St(\Phi, A[X], M[X]) \to \St(A[X, X^{-1}]))
\end{align}
Notice that for $ g\in G_+$ the element $ev_{X=0}(g)^{-1}g$ lies in $G_+^0$.

\begin{rem} Our goal is to rewrite $x_\alpha(aX^{-1}) \cdot z_\beta(f_1(X), f_2(X))$, where $f_1(X) \in M[X]$, $f_2(X) \in A[X]$ as
 $g \cdot x_\alpha(\frac{aX^{-1}}{1+am}) \cdot \{ 1 + am, X \}$ for some $g \in G_+$
 and $m \in M$. At first assume that $\Phi$ is simply-laced. \end{rem}
\begin{enumerate}
 \item {\it Case $\alpha \pm \beta \not \in \Phi \cup \{ 0\}$.} (Is this the same thing as $\alpha \perp \beta$?)
  Obviously, in this case $g = z_\beta(f_1(X), f_2(X))$, $m=0$.
 \begin{comment} 
 \item {\it Case $\alpha + \beta \in \Phi$, $f_1(X) = Xf_1'(X)$, $f_2(X) = 0$.}
 \[ x_\alpha(aX^{-1}) \cdot x_\beta(Xf_1'(X)) = x_{\alpha+\beta}(af_1'(X))x_\beta(Xf_1'(X)) x_\alpha(aX^{-1}) \]
 \end{comment}
 \item {\it Case $\alpha + \beta \in \Phi$, $f_1(X) = Xf_1'(X)$.}
 Start from this formula (centrality of $K_2$-paper):
 \begin{multline}
 x_\alpha(aX^{-1}) \cdot z_\beta(f_1(X), f_2(X)) = \\ = x_\alpha(af_1'(X)f_2(X)) \cdot x_{\alpha+\beta}(-N_{\beta, \alpha} af_1'(X)) \cdot z_\beta(f_1(X), f_2(X)) \cdot x_\alpha(aX^{-1}); 
 \end{multline}
 \item {\it Case $\beta - \alpha \in \Phi$, $f_1(X) = Xf_1'(X)$.}
 \begin{multline}
  x_{\alpha}(aX^{-1}) \cdot z_\beta(f_1(X), f_2(X)) = \\ = x_{\alpha}(-af_1'(X)f_2(X))\cdot x_{\alpha-\beta}(-N_{\beta,-\alpha} af_1'(X)f_2(X)^2) \cdot z_\beta(f_1(X), f_2(X)) \cdot x_{\alpha}(aX^{-1});
 \end{multline}
 \item {\it Case $\beta = -\alpha$, $f_1(X) = mX$, $f_2(X)=0$}. %CASE 1
For any $\alpha \in \Phi$ one can choose $\gamma$ such that $\langle \alpha, \gamma \rangle = -1$
(except in the sole case when $\alpha$ is a long root of a symplectic root system).
Now direct computation using~\eqref{eq:dennis-stein} and centrality of symbols shows that:
\begin{multline}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(X m) %= x_\alpha(X^{\langle \alpha, \gamma \rangle}a) x_{-\alpha}(X^{-\langle \alpha, \gamma \rangle}m) 
 = {}^{h_\gamma(X)}(x_\alpha(a) x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\frac{m}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\frac{a}{1+am}\right) \right) = \\
 x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha(X^{-1})^{-1} \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot \{1+am, X^{-1}\}^{-1} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}.
\end{multline}
 \item Essentially this allows one to cover the case $\beta = -\alpha$, $f_1(X) = Xm$ (with the same answer)
\end{enumerate}

\begin{lemma} Assume that $\Phi$ is simply-laced.
 For any $a\in A$, $\alpha\in \Phi$, $g_0 \in G^0_+$ there exist $m'\in m$ and $g'\in G_+$ such that 
 \[x_{\alpha}(aX^{-1}) g_0  = g' x_\alpha(-a(1+m')X^{-1}) \cdot \{X, 1+m'\}.\]
\end{lemma}
\begin{proof}
 Choose some parabolic subset $S$ of roots in $\Phi$ in such a way that its special part $\Sigma(S)$ does not contain the root $\alpha$.
 In view of~\cref{lem:Zgen} the subgroup $G_+^0$ is generated by images in $\St(\Phi, A[X, X^{-1}]$ of the elements of the set $\mathcal{Z}(\Sigma(S), A[X], Xm[X])$.
